---
title: "Data Visualization 3: The After Version"
description: |
  This plot updates the third data visualization by combining the two plots of Oregon into one and making it easier to read the text labels.
author:
  - name: Kathryn Denning
    url: https://psychology.uoregon.edu/profile/kdenning/
date: 03-19-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
#install.packages("rio")
#install.packages("here")
#install.packages("tidyverse")
#install.packages("magrittr")
#install.packages("RColorBrewer")
#install.packages("ggridges")
#install.packages("XML")
#install.packages("janitor")
#install.packages("colorspace")
#install.packages("ggrepel")
#install.packages("cowplot")
#install.packages("shadowtext")
library(rio)
library(here)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(ggridges)
library(XML)
library(janitor)
library(colorspace)
#library(ggsci)
library(ggrepel)
library(cowplot)
library(shadowtext)

knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

#read in data on all UFO citings in Jan 2019
x <- readHTMLTable("http://www.nuforc.org/webreports/ndxe201901.html")
#Turn it into a dataset
ufo_jan19 <- plyr::ldply(x, data.frame)
#power was out, so couldn't google how to add a key when there isn't one - so just wrote it into a csv and added in excel
write.csv(ufo_jan19,'ufo_jan19.csv')
ufo_jan19 <- read_csv("ufo_jan19.csv")

#cleaned dataset to be lowercase and to only include US states
clean <- ufo_jan19 %>%
  clean_names() %>% 
  filter(state != "AB") %>% 
  filter(state != "ON") %>% 
  filter(!is.na(state))

#Read in data of all ufo citings in Oregon ever
y <- readHTMLTable("http://www.nuforc.org/webreports/ndxlOR.html")
#Turn it into a dataset
ufo_oregon <- plyr::ldply(y, data.frame)

#cleaning the oregon dataset
ufo_oregon %<>% clean_names() 

#importing dataset with information about city, county, and population information
oregon_city_county <- import("oregon_city_county.csv")

#joining dataset of ufo data with oregon city and county data
oregon_join1 <- left_join(oregon_city_county, ufo_oregon)

#cleaning new dataset
oregon_join1 %<>% 
  na.omit() %>% 
  select(-id)

#getting informtion on the lat and long of each county for plotting
county <- ggplot2::map_data("county")

#changing the variables in county to have a key that will match ufo_oregon dataset
oregon_county <- county %>% 
  filter(region == "oregon") %>% 
  mutate(county = subregion)

oregon_join1$county <- tolower(oregon_join1$county)

#merging the county lat long dataset with the oregon_join1 dataset
oregon_join2 <- left_join(oregon_join1, oregon_county)

#getting the frequency of observations per county
oregon_plot_freq <- oregon_join2 %>% 
  group_by(county) %>% 
  count()

#joining the frequency back with the main dataset
oregon_join3 <- left_join(oregon_join2, oregon_plot_freq)

#dropping region and subregion from new overall dataset
oregon_join3 %<>% 
  select(-c(region, subregion))
state <- ggplot2::map_data("state")

#need to have a varaible with the full state names in lowercase to merge datasets
state_fullname <- datasets::state.name #calling dataset with state names
state_fullname <- sapply(state.name, tolower) #making them lowercase

state_abbrv <- datasets::state.abb #calling dataset with abbreviations to use as key when merging dataset to get full names in main dataset so that it can be combined with dataset used to make map of US

state_key <- data.frame(
  state_full = state_fullname,
  state_abrv = state_abbrv
) #making dataset with the two character strings state_fullname and state_abbrv

state_key %<>% clean_names

clean %<>% 
  mutate(state_abrv = state)

#creating the dataset for the state base for the plot

state_fixed <- state %>% 
  mutate(state_full = region) %>% 
  select(-region) %>% 
  clean_names

#making the base of oregon for the plot
oregon_base <- state_fixed %>% 
  filter(state_full == "oregon")

oregon_base <- ggplot(data = oregon_base, 
                     mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "white", fill = "white")

#code that removes axes from plot of US
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

#code that fills oregon with frequency data
oregon_filled <- oregon_base + 
  geom_polygon(data = oregon_join3, aes(fill = n), color = "white") +
      geom_polygon(color = "white", fill = NA) +
      scale_fill_continuous_sequential("viridis", name = "Frequency") +
  ditch_the_axes

#adding title
plot_title_freq <- oregon_filled + 
  labs(title = "Frequency of UFO citings per county in Oregon") +
  theme(plot.title = element_text(size = 18))

#figuring out top 5
oregon_join3 %>% 
  select(county, n, group) %>% 
  arrange(-n) %>% 
  unique() %>% 
  filter(n > 6693)

#creating dataset with names of counties and locations
freq_top5names <- c("Multnomah", "Lane", "Washington", "Marion", "Clackamas")
freq_y <- c(45.5, 44, 45.6, 44.8, 45.1)
freq_x <- c(-121.9, -122.5, -123.4, -122.1, -122)

freq_top5 <- data.frame(
  names = freq_top5names,
  y = freq_y,
  x = freq_x
)

#making a dataset for city, but couldnt' get the code later to run (it timed out...)
city <- c("Eugene", "Portland")
lat <- c(44.0521, 45.5122)
long <- c(-123.0868, -122.6587)

cities <- data.frame(
  city = city,
  long = long,
  lat = lat
)

#Adding labels for the top 5 counties and big cities
or_plot1 <- plot_title_freq + 
  geom_shadowtext(data = freq_top5, aes(label = names, x = x, y = y), inherit.aes = FALSE, size = 4, bg.colour='violetred4') 

or_plot1_cities <- or_plot1 + geom_point(aes(x = long, y = lat), data = cities, inherit.aes = FALSE, color = "red") + geom_text_repel(aes(x = long, y = lat, label = city), data = cities, inherit.aes = FALSE, color = "red", size = 4, fontface = "bold")

or_plot1_cities 
```


```{r chloropleth of proportion in oregon ever, include = FALSE}
#finding the proportion of observations per county
oregon_plot_proportion <- oregon_join3 %>% 
  group_by(county) %>% 
  mutate(freq_bypop = (n/county_pop)) %>% 
  summarize(freq_bypop_percounty = mean(freq_bypop))

#adding proportion back into overall dataset
oregon_join4 <- left_join(oregon_join3, oregon_plot_proportion)

#filling in base with county proportion information
oregon_filled_pro <- oregon_base + 
  geom_polygon(data = oregon_join4, aes(fill = freq_bypop_percounty), color = "white") +
      geom_polygon(color = "white", fill = NA) +
      scale_fill_continuous_sequential("viridis", name = "Proportion") +
  ditch_the_axes

#adding title
plot_title <- oregon_filled_pro + 
  labs(title = "Proportion of UFO citings per county in Oregon") +
  theme(plot.title = element_text(size = 18))

#Figuring out top 5 counties
county_names <- oregon_join4 %>% 
  select(county, freq_bypop_percounty, group) %>% 
  arrange(-freq_bypop_percounty) %>% 
  unique() %>% 
  filter(freq_bypop_percounty > 0.06523439)

#creating dataset with names of counties and locations
pro_top5names <- c("Baker", "Lane", "Curry", "Tillamook", "Lincoln")
pro_y <- c(44.7, 44, 45.5, 46, 44.7)
pro_x <- c(-117.7, -122.7, -124, -124.2, -124.2)

prop_top5 <- data.frame(
  names = pro_top5names,
  y = pro_y,
  x = pro_x
)

#Figure out total # of citings in OR ever
oregon_join4 %>% 
  count()

#Adding labels for the top 5 counties and Eugene/Portland
or_plot2 <- plot_title  + 
  geom_shadowtext(data = prop_top5, aes(label = names, x = x, y = y), inherit.aes = FALSE, size = 4, bg.colour='violetred4')

or_plot2_cities <- or_plot2 + geom_point(aes(x = long, y = lat), data = cities, inherit.aes = FALSE, color = "red") + geom_text_repel(aes(x = long, y = lat, label = city), data = cities, inherit.aes = FALSE, color = "red", size = 4, fontface = "bold")

#combining plots 1A and 1B into one plot side by side
combined2 <- plot_grid(or_plot1_cities, or_plot2_cities, labels = c("", ""), align = "hv", nrow = 2, hjust = -.05)

#Adding a caption
OR_final_withcounties <- ggdraw(add_sub(combined2, "Top five counties labeled. 121,456 documented citings statewide since 1905.\nData from National UFO Reporting Center and US Census Bureau.", x = .1, hjust = .1, vjust = 0, fontfamily = "",
  fontface = "plain", colour = "black", size = 10, angle = 0,
  lineheight = 0.9))
```

```{r fig.width=10, fig.height=10, echo = FALSE}
OR_final_withcounties
```

### What you're seeing:
This plot aids in the comparison of the two plots, demonstrating more clearly that the counties with the highest frequency of observations aren't generally the highest after taking into account county population size. However, Lane County is still in the top 5 counties after taking into account population size. The other counties with the highest proportion of observations are either along the coast or in the more remote and rural county of Baker County, which actually has the highest proportion of observations.

### Data visualization changes:
The plots were combined using the package "cowplot" and then the text labels for the top 5 counties were given an outline using the package "Shadowtext." The cities of Eugene and Portland were also added in order to orient people to the locations of the two most populous cities in Oregon.